The problem and promise of scale in multilayer animal social networks

### 2020-02-14 paper/manuscript.Rmd
reflow
hab perspective wg edits

## Habitat selection and space use
<!-- TODO: Gerard ‘openness’ paper -->
Social structure is influenced by resource availability and seasonality
[@Peignier_2019], configuration [@He_2019] and fragmentation [@Banks_2007].
Habitat selection, the disproportionate use of available resources or landscape
features by animals [@Manly_1993; @Boyce_1999], is a scale-dependent process
[@Mayor_2009]. Combined, social structure and habitat selection are intimately
linked as individuals must share space to interact and the distribution of
resources influences group dynamics and home range size [@Webber_2018].
<!--Animals select habitats at multiple scales ... (Mayor)  Mayor scale -->
Multilayer networks represent a novel framework for considering the influence of
habitat selection and space use on sociality across scales.

@@

<!-- TODO: point to figure 1 -->
Habitat selection depends on the scale at which it is measured and a singular
scale is insufficient for generalizing across scales [@Mayor_2009]. Habitat
selection is measured using resource selection functions (RSFs), based on the
proportion of used and available resources for an animal [@Manly_1993;
@Boyce_1999]. Johnson (1980) identified four orders of habitat selection: (1)
the geographical range of a species, (2) the home range of an individual, (3)
the habitat patches within the home range and (4) specific resources within a
habitat patch. Moving past this hierarchical perspective, Mayor (2009)
reviewed types of scale relevant in measurement of habitat selection and recommend using a continuum of scales for identifying characteristic scales of selection. For example, @Mayor_2007 showed caribou selection for shallow snow across all sales and used to a continuum of scales to avoid bias from narrowing the measurements to scales assumed to be associated with certain behaviors.

@@

The emergent relationship between habitat and sociality has been explored by the
distribution and phenology of resources and use of social information. Cougars
had increased home range overlap in regions with high hunting opportunity, a
proxy for resource richness [@Elbroch_2016]. Caribou showed seasonal differences
in sociality, where social associations were high in winter and contrastingly
low in summer when resource availability is homogenous [@Peigner_2019]. Bison
(*Bison bison*) showed stronger habitat selection when group size was larger and
avoidance of predation risk was a main factor contributing to their habitat
selection [@Fortin_2009]. These examples are implicitly multilayered, comparing sociality across contexts of seasonality or spatial context. <!--Despite these, *missing multilayer approach*.-->

@@

Multilayer networks can be used to explicitly consider habitat selection and sociality across scales. Animal social systems can be parsed by spatial contexts, defined, for example, by habitat types as in Figure \@ref(fig:lcres). The influence of resource phenology can be directly measured by differences in habitat selection and sociality across temporal layers. Habitat selection can be explicitly integrated using network layers defined by similarity of selection coefficients between individuals. These explicit integrations of habitat selection and sociality across social, spatial and temporal scales highlight the potential for novel insights from multilayer networks.

@@

Multilayer networks are complicated in that they can incorporate data as varied
as social interactions, landscape features, and seasons, all of which exist at
different scales and units of measurement. Our analysis of caribou multilayer
networks incorporated high resolution GPS telemetry data and spatial landcover
data is grounded in the context of observed animal social networks, as the
scales we chose to explore, e.g., number of observations, sociality across space
and time, are ones that can be observed as has traditionally been done in ethology. Thus, we
highlight ways in which scale ought to be considered for both observed and
remotely-sensed data when building social networks. Furthermore, GPS telemetry
data provides a link between large scale landscape properties and fine scale
social interactions as animal movement occurs across the landscape, and when
taken into account with multiple individuals can indicate association patterns.
Thus, a layer incorporating relocation data could be a great boon linking
spatial and social layer in multilayer networks.
<!-- JWT: I'm having a hard time getting to where I want to go, but basically I want to highlight that using GPS relocation data is reflective of more traditional observed networks like they wanted when they accepted our abstract, and this provides opportunities for social networks, particularly multilayer networks. It may be easier to scale relocation data and other landscape variables and can link to smaller scale observed interactions, providing a bridge between different types of data. -->

@@

1. How do social phenotypes vary across social, temporal, and spatial scales and
at which scale might these influence fitness? Are some scales better predictors for
different fitness metrics?
2. Why do individuals associate more strongly with certain conspecifics in some
habitats? What, if any, is the role of movement in the context of habitat
networks?
3. How, and why, does the spatial scale of perception influence, e.g. auditory,
visual, or chemotaxis, influence the  social environment? What is the role of
memory, and by extension social and spatial cognition, as animals navigate their
environment and make decisions about where to move next?

>>>>

## Habitat selection and space use
<!-- TODO: Gerard ‘openness’ paper -->
Social structure is influenced by resource availability and seasonality
[@Peignier_2019], configuration [@He_2019] and fragmentation [@Banks_2007].
Habitat selection, the disproportionate use of available resources or landscape
features by animals [@Manly_1993; @Boyce_1999], is a scale-dependent process
[@Mayor_2009]. Social structure and habitat selection are intimately
linked: as individuals must share space to interact, and the distribution of
resources influences group dynamics and home range size [@Webber_2018].
<!--Animals select habitats at multiple scales ... (Mayor)  Mayor scale -->
Multilayer networks represent a novel framework for considering the influence of
habitat selection and space use on sociality across scales.

@@

<!-- TODO: point to figure 1 -->
Habitat selection depends on the scale at which it is measured and a singular
scale is insufficient for generalizing across scales [@Mayor_2009]. Habitat
selection is measured using resource selection functions (RSFs), based on the
proportion of used and available resources for an animal [@Manly_1993;
@Boyce_1999]. Johnson (1980) identified four orders of habitat selection: (1)
the geographical range of a species, (2) the home range of an individual, (3)
the habitat patches within the home range and (4) specific resources within a
habitat patch. Moving past this hierarchical perspective, Mayor (2009) reviewed
types of scale relevant in measurement of habitat selection and recommend using
a continuum of scales for identifying characteristic scales of selection. For
example, @Mayor_2007 showed caribou select for shallow snow at all scales
and by using a continuum of scales, avoided the bias of predefined scales assumed to be associated with certain behaviors.

@@

The emergent relationship between habitat and sociality has been explored by the
distribution and phenology of resources and use of social information. Cougar
home ranges overlapped more in regions with high hunting opportunity, a
proxy for resource richness [@Elbroch_2016]. Caribou sociality differed between seasons, where social associations were stronger in winter and contrastingly
low in summer when resource availability are more homogenous [@Peigner_2019]. Bison
(*Bison bison*) showed stronger habitat selection in larger groups and
avoidance of predation risk was a main factor contributing to their habitat
selection [@Fortin_2009]. These examples, comparing sociality and habitat selection across temporal or spatial contexts, are implicitly multilayer analyses. <!--Despite these, *missing multilayer approach*.-->

@@

Multilayer networks explicitly consider habitat selection and
sociality across scales. Animal social systems can be parsed by spatial
contexts, defined, for example, by habitat types as in Figure \@ref(fig:lcres).
The effect of resource phenology can be directly measured by differences in
habitat selection and sociality across temporal layers. Habitat selection can be
explicitly integrated using network layers defined by similarity of selection
coefficients between individuals. These explicit integrations of habitat
selection and sociality across social, spatial and temporal scales highlight the
potential for novel insights from multilayer networks.

@@

Multilayer networks are complicated in that they can incorporate data as varied
as social interactions, landscape features, and seasons, all of which exist at
different scales and units of measurement. Our analysis of caribou multilayer
networks incorporated high resolution GPS telemetry data and spatial landcover
data is grounded in the context of observed animal social networks, as the
scales we chose to explore, e.g., number of observations, sociality across space
and time, are ones that can be observed as has traditionally been done in
ethology. Thus, we highlight ways in which scale ought to be considered for both
observed and remotely-sensed data when building social networks. Furthermore,
GPS telemetry data provides a link between large scale landscape properties and
fine scale social interactions as animal movement occurs across the landscape,
and when taken into account with multiple individuals can indicate association
patterns. Thus, a layer incorporating relocation data could be a great boon
linking spatial and social layer in multilayer networks.

<!-- JWT: I'm having a hard time getting to where I want to go, but basically I want to highlight that using GPS relocation data is reflective of more traditional observed networks like they wanted when they accepted our abstract, and this provides opportunities for social networks, particularly multilayer networks. It may be easier to scale relocation data and other landscape variables and can link to smaller scale observed interactions, providing a bridge between different types of data. -->

@@

1. How do social phenotypes vary across social, temporal, and spatial scales and
at which scale might these influence fitness? Are some scales better predictors
for different fitness metrics?

2. Why do individuals associate more strongly
with certain conspecifics in some habitats? What, if any, is the role of
movement in the context of habitat networks?

3. How, and why, does the spatial
scale of perception influence, e.g. auditory, visual, or chemotaxis, influence
the  social environment? What is the role of memory, and by extension social and
spatial cognition, as animals navigate their environment and make decisions
about where to move next?
